### 2015-06-27 text/ms.Rmd
Editing throughout
Including switching to numbered references

@@

@@

Ensemble methods, which combine the output from multiple models as predictors in a new 'ensemble' model, are one solution to this problem. A simple ensemble model might take the average of multiple model outputs, perhaps weighted by some goodness of fit. A complex ensemble model might also take into account non-linear interactions between model outputs and additional covariates. Ensemble models are popular in other fields including public health [e.g. @thomson2006], agriculture [e.g. @cantelaube2005], climate science [e.g. @murphy2004; @tebaldi2007; @pierce2009], and machine learning [e.g. @dietterich2000]. We think they have tremendous opportunity in ecological resource management.

In fisheries science, a common task is estimating the status and trajectory of an exploited fish population. For the majority of fish stocks, we have limited biological information and stock status isn't known [@fao2014]. In recent years, a number of methods have been proposed to derive population status based on the limited information that is broadly available: fisheries catch time series and basic knowledge about the productivity of the species. Recently, @rosenberg2014 investigated the ability for these models to estimate population status through a large-scale simulation model. A key finding was that these models frequently disagree on population status and any one model has relatively poor predictive ability on average (e.g. Fig. \ref{motivate}).

@@

We tested the ability of ensemble models to improve estimates of population status and trajectory when applied to both a large simulated dataset and a global database of assessed stock status. Here we describe the datasets, individual models of population status, and ensemble models to combine those estimates. We then describe how we evaluated the ability of the various models to estimate population status.

@@

@rosenberg2014 ran ten iterations for each combination of factors adding stochastic draws of recruitment and catch-recording variability each time to generate a total of 5760 stocks. The simulation models were built in the package FLR [@kell2007] for the statistical software \textsf{R} [@r2015]. @rosenberg2014 provide a full description of the simulation model and the code to generate the simulations is available at <https://github.com/flr/StockSims> (TODO: not there).

@@

We fit four individual data-limited models to estimate \bbmsy\\. Three of the models are mechanistic and based generally on @schaefer1954 biomass dynamics of the form

@@

\noindent
where $\hat{B}_{t+1}$ represents predicted biomass at time $t$ plus one year, $B_t$ represents biomass at time $t$, $r$ represents the intrinsic population growth rate, $B_0$ represents unfished biomass or carrying capacity, and $C$ represents catch. @rosenberg2014 provide a full summary of these four methods and code to fit all the models is available in an accompanying package **datalimited** for the statistical software \textsf{R}. In summary:

* *COM-SIR* (catch-only-model with sample-importance-resampling) is a coupled
  harvest-dynamics model developed by @vasconcellos2005. Biomass is assumed to
  follow a Schaefer model and harvest dynamics are assumed to follow a logistic
  model. The model is fit with a sample-importance-sampling algorithm developed
  in @rosenberg2014.

@@

* *SSCOM* (state-space catch-only-model) TODO: Jim -- can you give a ~2 sentence summary?

* *mPRM* (modified panel regression model) is a modified version of the
  panel-regression model used in @costello2012. Unlike the other models
  considered, mPRM is empirical and not mechanistic --- it uses the RAM Legacy
  Stock Assessment database to fit a linear model to a series of
  characteristics of the catch time series and stock with stock-assessed
  \bbmsy\\ as the response. The model used in this paper and @rosenberg2014 is
  modified from the original in that it condenses the life-history categories
  into three categories to match the simulated dataset.
  <!--TODO: what are we doing with the RAM stocks again?-->
  <!--TODO: bias adjustment?-->
  <!--TODO: build in fig:didactic -->

@@

Ensemble methods allow us to incorporate additional covariates into our models and potentially leverage interactions between these covariates and individual model predictions. Additional covariates could be, for example, life-history characteristics, information on exploitation patterns, or statistical properties of the data. For simplicity, and to allow us to apply models developed with the simulated dataset to the real-world dataset, we added only one set of additional covariates: spectral properties of the catch time series. Spectral analysis decomposes a time series into the frequency domain and provides a means of describing the statistical properties of the catch series that is independent of time series length and magnitude of catch. We fit spectral models to the scaled catch time series (catch divided by maximum catch) with the `spec.ar` function in \textsf{R} and recorded representative short- and long-term spectral densities of 0.05 and 0.2, which correspond to 20- and 5-year cycles.

@@

The individual models we seek to combine with ensemble models provide time series of \bbmsy\\. Therefore, we can use ensemble models to estimate multiple properties of these status time series. Here, we focus on two values: the mean and slope of \bbmsy\\ in the last 5 years. Together, these quantities address the recent state and trajectory of status, both of which may be of management and conservation interest. To avoid undue influence of the end points of the time series on the calculated slope, we measured the slope as the Theil-Sen estimator of median slope [@theil1950].

@@

We calculated the average ensemble models as the geometric mean of the individual model \bbmsy\\ estimates and the mean of the individual model \bbmsy\\ slopes. We fit the linear model ensemble with second-order interactions. We fit two machine learning ensemble models based on regression trees. Regression trees sequentially determine what value of a predictor best splits the response data into two 'branches' based on a specified loss function. In random forests, a series of regression trees are built on a random subset of the data and random subset of the covariates of the model (REF). In generalized boosted models (GBMs), each subsequent model is fit to the residuals from the previous model; data points that are fit poorly in a given model are given more weight in the next model [@elith2008]. Random forests and GBMs can provide strong predictive performance and fit highly non-linear relationships [e.g. @elith2008; @hastie2009]. We fit random forest models with the **randomForest** package [@liaw2002] for \textsf{R} with the default argument values. We fit boosted regression tree models with the **GBM** package [@ridgeway2015] for \textsf{R}. Based on cross-validation with the **caret** \textsf{R} package, we fit GBMs with 2000 trees, an interaction depth of $6$, a learning rate (shrinkage parameter) of $0.01$, and all other arguments at their default values.

@@

Certain conditions will make some ensemble models more effective than others.
First, ensembles will be most effective when they are comprised of diverse
individual models that choose different structural model forms, explore
contrasting but plausible ranges of parameter values, or make errors in
uncorrelated ways [e.g. @ali1996; @dietterich2000; @tebaldi2007]. Such
individual models would be expected to perform well in different conditions and
an ensemble model can exploit the best predictive performance of each. Second,
ensemble models will be most effective when they are not overfit to the
training dataset. Cross-validation testing [e.g. @caruana2004; @hastie2009] and
methods that are robust to over-fitting such as random forests [@breiman2001],
may help avoid overfitting ensemble models. We note that our simplest ensemble
model, an average of individual model predictions, performed approximately as
well as complex machine learning models when we applied our ensembles to an
entirely new dataset (Fig. \ref{performance}b) and averages are a common
ensemble approach in climate science [@ipcc2013]. Third, ensemble
models will be most effective when they are trained on data that are
representative of the dataset of interest [e.g. @knutti2009; @weigel2010].
Cross-validation within a training dataset will provide a optimistically biased
impression of predictive performance if the training dataset fundamentally
differs from the dataset of interest [@hastie2009].

Multi-model inference in the form of coefficient averaging weighted by
information theoretics such as the Akaike Information Criteron (AIC) is a
common analytical approach in ecology [e.g. @burnham2002; @grueber2011]. The
ensemble methods described in this paper share similarities with coefficient
averaging but differ in other important ways. Ensemble models and coefficient
averaging share the long-held notion that multiple working hypotheses can
contribute useful information for inference [@chamberlin1890]. A fundamental
difference is that coefficient averaging focuses on a averaging *coefficients*
whereas ensembles instead average *predictions*. By focussing on predictions,
ensembles do not suffer from the muddled inference of coefficient averaging
when coefficients covary [@cade2015] and are a more general purpose tool: they
do not require information theoretics, they can combine different types of
models (e.g. parametric and non-parametric models or frequentist and
Bayesian predictions), and they can combine predictions in complex non-linear
ways that may depend on additional covariates.

@@

Combining predictions from multiple models via ensemble methods is also broadly
useful in other subfields of ecology and fisheries science. Predictions about
extinction risk are widely used at the national (e.g. the US Endangered Species
Act and the Canadian Species at Risk Act) and international [e.g. the IUCN Red
List, @iucn2015] levels. These risk assessments generally involve fitting
regression models to outcomes for individual species along with predictors of
species risk [e.g. @anderson2011a;@pinsky2011], or population dynamics models to
data for individual species [e.g. @dfo2010]. Both types of models are prone to
error caused by model-misspecification and therefore results are sensitive to
decisions about model structure [@brooks2015]. Although there are options to
account for potential model-misspecification in determination of species risk
[e.g., coefficient averaging, @burnham2002; or semi-parametric methods,
@thorson2014], ensemble methods are a relatively simple way to combine
predictions in a transparent manner. Beyond estimates of status and trajectory,
ensemble methods could be used to increase the robustness of spatial
predictions when designing networks of protected areas [@rassweiler2014], or to
forecast potential spatial shifts in species distribution given climate impacts
[@harsch2014].

@@

@@

>>>>

<!--200 words-->

@@

<!-- 3-6 keywords -->
Keywords: data-limited fisheries, model averaging, population dynamics, ... *(up to 6)*

@@

Ensemble methods, which combine the output from multiple models as predictors
in a new 'ensemble' model, are one solution to this problem. A simple ensemble
model might take the average of multiple model outputs, perhaps weighted by
some goodness of fit. A complex ensemble model might also take into account
non-linear interactions between model outputs and additional covariates.
Ensemble models are popular in other fields including public health
[@thomson2006], agriculture [@cantelaube2005], climate science [@murphy2004;
@tebaldi2007; @pierce2009], and machine learning [@dietterich2000]. We think
they have tremendous opportunity in ecological resource management.

In fisheries science, a common task is estimating the status and trajectory of
an exploited fish population. For the majority of fish stocks, we have limited
fisheries independent data to estimate stock status [@fao2014]. In recent
years, a number of methods have been proposed to derive population status based
on the limited information that is broadly available: fisheries catch time
series and basic knowledge about the productivity of the species (e.g.
[@vasconcellos2005; @martell2013]). Recently, Rosenberg et al. [@rosenberg2014]
investigated the ability for four such models to estimate population status through
a large-scale simulation experiment. A key finding was that these models
frequently disagree on population status and any one model has relatively poor
predictive ability on average (e.g. Fig. \ref{motivate}).

@@

We tested the ability of ensemble models to improve estimates of population
status and trajectory when applied to both a large simulated dataset and a
global database of assessed stock status. Here we describe the datasets,
individual models of population status, and ensemble methods to combine those
estimates. We then describe how we evaluated the ability of the various models
to estimate population status.

@@

Rosenberg et al. [@rosenberg2014] ran ten iterations for each combination of
factors adding stochastic draws of recruitment and catch-recording variability
each time to generate a total of 5760 stocks. The simulation models were built
in the package `FLR` [@kell2007] for the statistical software \textsf{R}
[@r2015]. The simulation model and the code to generate the simulations is
available at <https://github.com/flr/StockSims> (TODO: not there).

@@

We fit four individual data-limited models to estimate \bbmsy\\. Three of the
models are mechanistic and based generally on Schaefer biomass dynamics
[@schaefer1954] of the form

@@

\noindent
where $\hat{B}_{t+1}$ represents predicted biomass at time $t$ plus one year,
$B_t$ represents biomass at time $t$, $r$ represents the intrinsic population
growth rate, $B_0$ represents unfished biomass or carrying capacity, and $C$
represents catch. @rosenberg2014 provide a full summary of these four methods
and code to fit all the models is available in an accompanying package
`datalimited` for the statistical software \textsf{R}. In summary:

* *COM-SIR* (catch-only-model with sample-importance-resampling) is a coupled
  harvest-dynamics model @vasconcellos2005. Biomass is assumed to follow a
  Schaefer model and harvest dynamics are assumed to follow a logistic model.
  The model is fit with a sample-importance-sampling algorithm developed in
  @rosenberg2014.

@@

* *SSCOM* (state-space catch-only model) TODO: Jim -- can you give a ~2 sentence summary?

* *mPRM* (modified panel regression model) is a modified version of the
  panel-regression model from Costello et al. [@costello2012]. Unlike the other
  models considered, mPRM is empirical and not mechanistic---it uses the RAM
  Legacy Stock Assessment database to fit a linear model to a series of
  characteristics of the catch time series and stock with stock-assessed
  \bbmsy\\ as the response. The model used in this paper and @rosenberg2014 is
  modified from the original in that it condenses the life-history categories
  into three categories to match the simulated dataset, removes the maximum
  catch predictor since the absolute catch in the simulated dataset is
  arbitrary, and does not implement the bias correction needed in Costello et
  al. [@costello2012] for deriving aggregate estimates of mean status across
  multiple stocks. (*TODO: can we avoid listing most of these modifications or
  relegate them to the supplement?*)

  <!--TODO: build in fig:didactic -->

@@

Ensemble methods allow us to incorporate additional covariates into our models
and potentially leverage interactions between these covariates and individual
model predictions. Additional covariates could be, for example, life-history
characteristics, information on exploitation patterns, or statistical
properties of the data. For simplicity, and to allow us to apply models
developed with the simulated dataset to the real-world dataset, we added one
set of additional covariates: spectral properties of the catch time series.
Spectral analysis decomposes a time series into the frequency domain and
provides a means of describing the statistical properties of the catch series
that is independent of time series length and magnitude of catch. We fit
spectral models to the scaled catch time series (catch divided by maximum
catch) with the `spec.ar` function in \textsf{R} and recorded representative
short- and long-term spectral densities at frequencies of 0.20 and 0.05, which
correspond to 5- and 20-year cycles.

@@

The individual models we seek to combine with ensemble models provide time
series of \bbmsy\\. Therefore, we can use ensemble models to estimate multiple
properties of these status time series. Here, we focus on two properties: the
mean and slope of \bbmsy\\ in the last 5 years. Together, these quantities
address the recent state and trajectory of status, which are both of management
and conservation interest (e.g. [@mace2008]). To avoid undue influence of the
end points of the time series on the calculated slope, we measured the slope as
the Theil-Sen estimator of median slope [@theil1950].

@@

We calculated the average ensemble models as the geometric mean of the
individual model \bbmsy\\ estimates and the mean of the individual model
\bbmsy\\ slopes. We fit the linear model ensemble with second-order
interactions. We fit two machine learning ensemble models based on regression
trees. Regression trees sequentially determine what value of a predictor best
splits the response data into two 'branches' based on a specified loss
function. In random forests, a series of regression trees are built on a random
subset of the data and random subset of the covariates of the model (REF). In
generalized boosted models (GBMs), each subsequent model is fit to the
residuals from the previous model; data points that are fit poorly in a given
model are given more weight in the next model [@elith2008]. Random forests and
GBMs can provide strong predictive performance and fit highly non-linear
relationships [@elith2008; @hastie2009]. We fit random forest models with the
`randomForest` package [@liaw2002] for \textsf{R} with the default argument
values. We fit boosted regression tree models with the `gbm` package
[@ridgeway2015] for \textsf{R}. Based on cross-validation with the **caret**
\textsf{R} package, we fit GBMs with 2000 trees, an interaction depth of $6$, a
learning rate (shrinkage parameter) of $0.01$, and all other arguments at their
default values.

@@

Certain conditions will make some ensemble models more effective than others.
First, ensembles will be most effective when they are comprised of diverse
individual models that choose different structural model forms, explore
contrasting but plausible ranges of parameter values, or make errors in
uncorrelated ways [@ali1996; @dietterich2000; @tebaldi2007]. Such
individual models would be expected to perform well in different conditions and
an ensemble model can exploit the best predictive performance of each. Second,
ensemble models will be most effective when they are not overfit to the
training dataset. Cross-validation testing [@caruana2004; @hastie2009] and
methods that are robust to over-fitting such as random forests [@breiman2001],
may help avoid overfitting ensemble models. We note that our simplest ensemble
model, an average of individual model predictions, performed approximately as
well as complex machine learning models when we applied our ensembles to an
entirely new dataset (Fig. \ref{performance}b) and averages are a common
ensemble approach in climate science [@ipcc2013]. Third, ensemble
models will be most effective when they are trained on data that are
representative of the dataset of interest [@knutti2009; @weigel2010].
Cross-validation within a training dataset will provide a optimistically biased
impression of predictive performance if the training dataset fundamentally
differs from the dataset of interest [@hastie2009].

Multi-model inference in the form of coefficient averaging weighted by
information theoretics such as the Akaike Information Criteron (AIC) is a
common analytical approach in ecology [@burnham2002; @grueber2011]. The
ensemble methods described in this paper share similarities with coefficient
averaging but differ in other important ways. Ensemble models and coefficient
averaging share the long-held notion that multiple working hypotheses can
contribute useful information for inference [@chamberlin1890]. A fundamental
difference is that coefficient averaging focuses on a averaging *coefficients*
whereas ensembles instead average *predictions*. By focussing on predictions,
ensembles do not suffer from the muddled inference of coefficient averaging
when coefficients covary [@cade2015] and are a more general purpose tool: they
do not require information theoretics, they can combine different types of
models (e.g. parametric and non-parametric models or frequentist and
Bayesian predictions), and they can combine predictions in complex non-linear
ways that may depend on additional covariates.

@@

Combining predictions from multiple models via ensemble methods is also broadly
useful in other subfields of ecology and fisheries science. Predictions about
extinction risk are widely used at the national (e.g. the US Endangered Species
Act and the Canadian Species at Risk Act) and international (e.g. the IUCN Red
List, [@iucn2015]) levels. These risk assessments generally involve fitting
regression models to outcomes for individual species along with predictors of
species risk (e.g. [@anderson2011a; @pinsky2011)], or population dynamics models
to data for individual species (e.g. [@dfo2010]). Both types of models are prone
to error caused by model-misspecification and therefore results are sensitive
to decisions about model structure [@brooks2015]. Although there are options to
account for potential model-misspecification in determination of species risk
(e.g., coefficient averaging, [@burnham2002], or semi-parametric methods,
[@thorson2014]), ensemble methods are a relatively simple way to combine
predictions in a transparent manner. Beyond estimates of status and trajectory,
ensemble methods could be used to increase the robustness of spatial
predictions when designing networks of protected areas [@rassweiler2014], or to
forecast potential spatial shifts in species distribution given climate impacts
[@harsch2014].

@@

# Authors' contributions

TODO *Example: AB carried out the molecular lab work, participated in data
analysis, carried out sequence alignments, participated in the design of the
study and drafted the manuscript; CD carried out the statistical analyses; EF
collected field data; GH conceived of the study, designed the study,
coordinated the study and helped draft the manuscript. All authors gave final
approval for publication.*

@@

TODO

# Funding

TODO
